---
title: "R Visualization Portfolio Project"
author: "Sohaib"
date: "2023-11-28"
output: html_document
---

## Setting up my environment 
 Notes: setting up my environment by loading the 'tidyverse', 'palmer penguins', and 'diamonds' packages and data sets.
 
```{r loading packages}
library("ggplot2")
library("palmerpenguins")
```

## Viewing our Penguins Dataset
Here we will go through viewing our penguins data set

```{r ggplot for penguin data}
library(ggplot2)
library(palmerpenguins)
data(penguins)
View(penguins)
```

## Visualizations
Here We will go through a series of visualization

## Flipper and body mass 
Here, we plot flipper length against body mass

```{r ggplot for penguin data visualization}

ggplot(data = penguins) +

  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))

```

## Flipper and body mass by species
Here, we plot flipper length against body mass. we will color, change the shape and the size of the species

```{r}
ggplot(data=penguins) + geom_point(mapping = aes(x=flipper_length_mm,y=body_mass_g, color=species, shape=species, size=species))
```

## Bill length and bill depth by species
Here, we plot bill length against bill depth, then we will use alpha for species

```{r}
ggplot(data=penguins) + geom_point(mapping = aes(x=bill_length_mm,y=bill_depth_mm))
ggplot(data=penguins) + geom_point(mapping = aes(x=flipper_length_mm,y=body_mass_g, alpha=species))
```

## Flipper and body mass in orange
Here, we plot flipper length against body mass with coloring all the points

```{r}
ggplot(data=penguins) + geom_point(mapping = aes(x=flipper_length_mm,y=body_mass_g),color="orange")

```

## Flipper and body mass in a smooth way
Here, we plot flipper length against body mass in a smooth way

```{r}
ggplot(data=penguins) + geom_smooth(mapping = aes(x=flipper_length_mm,y=body_mass_g))
```

## Flipper and body mass two ways represntation

Here we will combine the points with smooth function

```{r}
ggplot(data=penguins) + geom_smooth(mapping = aes(x=flipper_length_mm,y=body_mass_g)) +
  geom_point(mapping = aes(x=flipper_length_mm,y=body_mass_g))
```

## Flipper and body mass by species in smooth way
Here, we plot flipper length against body mass and separating each species
```{r}
ggplot(data=penguins) + 
  geom_smooth(mapping = aes(x=flipper_length_mm,y=body_mass_g, linetype=species))
```

## Flipper and body mass by species by adding small variation
Here, we plot flipper length against body mass and add a small amount of random variation to the location of each point in a graph

```{r}
ggplot(data=penguins) + 
  geom_jitter(mapping = aes(x=flipper_length_mm,y=body_mass_g))
```

## Flipper and body mass by species and sex
Here, we plot flipper length against body mass and look at the breakdown by species and sex

```{r}
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point(aes(color=species,shape=species))+
  
  facet_wrap(~sex)
```

## Cut of the diamonds
Here, we are going to use diamonds data set with bar graphs, and look at the breakdown by cut.

```{r}

ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill=cut ))
```

## Cut of the diamonds
Here, we we plot cut breakdown by clarity

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill=clarity ))
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill=clarity )) + theme(axis.text.x = element_text(angle = 45))+ 
  facet_wrap(~cut)
```

